CudeleFS: Programmable Consistency and Durability\\in a Global Namespace

### 2017-04-02 paper/sections/4_methodology.tex
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Of the 6 mechanisms in
Figure~\ref{fig:decouple} only 4 had to implemented and just 1 required changes to
the underlying storage system itself.

\subsubsection{No Changes} ``RPCs" and ``Stream" are part of the default CephFS
implementation; they are used to get strong consistency and global durability,
respectively.  Using existing configuration settings in Ceph we can turn
``Stream" on and off.  If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run.

\subsubsection{Library} for ``Create", ``Nonvolatile Apply", ``Local/Global
Persist", CudeleFS provides a library for clients to link into and all
operations are performed by the client.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal.  For
``Local Persist", clients write serialized log events to a file on local disk
and for ``Global Persist", clients push the journal into the objects store. The
overheads for both ``Local Persist" and ``Global Persist" is the write
bandwidth of the local disk and object store, respectively.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  These implementations required no changes to CephFS
because the metadata servers know how to read the events the library is writing
into the object store.  By re-using the journal subsystem to implement the
namespace decoupling, Cudele leverages the write/read optimized data
structures, the formats for persisting events (similar to TableFS's
SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
onto the internal namespace data structures.

@@

% - how it makes no gaurantees
\subsubsection{Changes to Metadata Server} the ``Volatile Apply" mechanism
takes an in-memory journal on the client and applies the updates directly to
the in-memory namespace maintained by the metadata servers. We say volatile
because -- in exchange for peak performance -- Cudele makes no consistency or
durability guarantees while ``Volatile Apply" is executing.  If a concurrent
update from a client occurs there is no rule for resolving conflicts and if the
client or metadata server crashes there may be no way to recover.

% difference between apply and volatile apply
In contrast, ``Nonvolatile Apply" uses the the object store to apply the
journal of updates from the client onto the metadata servers' metadata store.
``Nonvolatile Apply" is safer but has a performance overhead because objects in
the metadata store need to be read from and written back to the object store.

@@

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. 
\label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

@@

%\begin{listing}
%\begin{minted}[frame=single,
%               framesep=3mm,
%               xleftmargin=21pt,
%               tabsize=4]{js}
%{     
%    "allocated_inodes": "-"
%    "interfere_policy": "-"
%    "consistency": "RPCs"
%    "durability": "stream"
%}
%\end{minted}
%\caption{Existing CephFS on Cudele.}
%\label{src:batchfs}
%\end{listing}

>>>>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisified by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes an in-memory journal on the
client and applies the updates directly to the in-memory namespace maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

% difference between apply and volatile apply
The biggest difference between ``Volatile Apply" and ``Nonvolatile Apply" is
the medium they use to communciate. ``Volatile Apply" applies updates directly
to the metadata servers' metadata store while ``Nonvolatile Apply" uses the
object store to communicate the journal of updates from the client to the
metadata servers.  ``Nonvolatile Apply" is safer but has a large performance
overhead because objects in the metadata store need to be read from and written
back to the object store.

\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
mechanisms used by default in CephFS.  Using existing configuration settings in
Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
will not save journals in the object store and the daemons that apply the
journal to the metadata store will never run.  For ``Local Persist", clients
write serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively.  These persist mechanisms are part a library that
links into the client.

@@

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}

\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

@@

%\begin{listing}
%\begin{minted}[frame=single,
%               framesep=3mm,
%               xleftmargin=21pt,
%               tabsize=4]{js}
%{     
%    "allocated_inodes": "-"
%    "interfere_policy": "-"
%    "consistency": "RPCs"
%    "durability": "stream"
%}
%\end{minted}
%\caption{Existing CephFS on Cudele.}
%\label{src:batchfs}

%\end{listing}

### 2017-04-02 paper/sections/5_implementation.tex
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
CudeleFS; namely, we try to leverage components inside Ceph to inherit the
robustness and correctness of the internal subsystems. In this section, we
first describe a Ceph internal subsystem or component and then we show how we
use it in CudeleFS.

@@

% how i used it
\noindent\textbf{Cudele}: uses the metadata store format to write objects
formatted like the journal into the object store. By writing with the same
format, the metadata servers can read and use the recovery code to materialize
the updates from a client's decoupled namespace ({\it i.e.} merge).

@@

\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
server to replay updates onto the in-memory metadata store.  When the clients
are ready to merge their updates back into the global namespace, they pass a
binary file of  metadata updates to the metadata server. The metadata server
uses its recovery code to ``replay" the updates.

@@

\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
journal in memory and append events to their own copy. When they are done they
export their journal to a binary file. Clients also use the journal tool to
save their updates to disk or to the object store.

@@

\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

>>>>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
CudeleFS; namely, we try to leverage components inside Ceph to inherit the
robustness and correctness of the internal subsystems. Using this `dirty-slate'
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a Ceph internal subsystem or
component and then we show how we use it in CudeleFS.

%- no change (rpcs, stream)s, library(create, nonvolatile apply, local/global
%persist),  change to metadata server (nonvolatile apply)

@@

% how i used it
\noindent\textbf{CudeleFS}: uses the metadata store format to write objects
formatted like the journal into the object store. By writing with the same
format, the metadata servers can read and use the recovery code to materialize
the updates from a client's decoupled namespace ({\it i.e.} merge).  These
implementations required no changes to CephFS because the metadata servers know
how to read the events the library is writing into the object store.  By
re-using the journal subsystem to implement the namespace decoupling, Cudele
leverages the write/read optimized data structures, the formats for persisting
events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
functions for replaying events onto the internal namespace data structures.\\

\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply

@@

\noindent\textbf{CudeleFS}: uses the journal subsystems within the metadata
server to replay updates onto the in-memory metadata store.  When the clients
are ready to merge their updates back into the global namespace, they pass a
binary file of  metadata updates to the metadata server. The metadata server
uses its recovery code to ``replay" the updates.\\

\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply, Local/Persist

@@

\noindent\textbf{CudeleFS}: using the journal tool, clients can materialize the
journal in memory and append events to their own copy. When they are done they
export their journal to a binary file. Clients also use the journal tool to
save their updates to disk or to the object store.\\

\noindent\textbf{Used to Implement}: Create, Volatile/Nonvolatile Apply

@@

\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.\\

\noindent\textbf{Used to Implement}: Nonvolatile Apply, Create

### 2017-04-02 paper/sections/6_results.tex
Fixes typos and minor grammar edits (#15)
partial
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsection{Benefits of Isolated Subtree Policies}
% setup: 2cs sep dirs, 2cs DN, 1c malicious write
In this section, we show the isolation benefits of subtree policies.  We repeat
the experiment from Figure~\ref{fig:interfere-b}, where clients write to their
own private directories and another client interferes at 30 seconds.  We also
have 2 clients write to decoupled namespaces and merge their updates 90
seconds. The

>>>>

\subsection{Benefits of Isolated Subtree Policies}

\begin{listing}[tb]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\caption{Existing CephFS on Cudele with blocking.}
\label{src:block}
\end{listing}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
Using the ``allow" and ``block" API, users can isolate directories from
interferring clients. Performance with blocking turned is the same as
``isolated" from Figure~\ref{fig:interfere-b}.
\label{fig:slowdown-allow-block}}
\end{figure}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how CudeleFS can be programmed to block interferring clients. We
use the same problematic workload from Figure~\ref{fig:interfere-b}, where
clients write to their own private directories and another client interferes at
30 seconds.  We also have another client write to a decoupled namespace and
merge its updates at 90 seconds.  We equip each client directory with the
configuration in Listing~\ref{src:block}.  Each directory functions with RPCs
and stream enabled -- which is the default implementation of CephFS. The only
difference is that we enable blocking on each subtree so while the tree behaves
like CephFS, all interferring operations will be returned with \texttt{-EBUSY}.
Note that IndexFS does a similar operation with leases except clients block.

% results
The benefits of this isolations are shown next to the bars from
Figure~\ref{fig:interfere-b} in Figure~\ref{fig:slowdown-allow-block}. Because
\{``allow", ``block"\} has the same performance as \{``interfere",
``isolated"\}, we draw the following three conclusions: (1) clients that use the
API to block interferring clients  get the same performance as isolated
clients, (2) there is a neglible effect on performance for the extra work the
metadata server does to return \texttt{-EBUSY}, and (3) merging updates from
the decoupled client has a neglible effect on performance.\\

\noindent\textbf{Takeaway}: the API is sufficient for isolating performance
when applications require it. This is a way of controlling consistency.
